\papertitle

### 2012-10-13 manuscript.tex
Fix oops
adits

%----------------------------------------------------------------------------%
% PACKAGES
\usepackage{color}
\usepackage[pdftex]{hyperref}
\usepackage{amsmath}

\newlength\eqnindent
\newlength\eqncolwid
\newcommand\eqntext[2][0pt]{%
  \setlength\parindent{0pt}%
      \setlength\eqnindent{#1}\setlength\eqncolwid{\linewidth}%
  \setlength\eqncolwid{\dimexpr\linewidth-\eqnindent}%
  \hspace*{\eqnindent}\parbox{\eqncolwid}{#2}%
}

@@

We also introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
features clap and fling wings, improves upon a previous design by 
utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a 
payload of 2.8 grams. We augment the ornithopter's built-in 
gyroscope-based control with a ground station, which has power and weight 
requirements appropriate for deployment on a millirobotic ground vehicle. The ground 
station provides heading estimates to the ornithoper by running a 
real-time motion tracking algorithm on a live video stream. This allows 
the ornithopter to reliably negotiate flight through narrow openings.

@@

Autonomous control of flapping winged MAVs is a complex undertaking because of
the non-linear flight dynamics as presented
in~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. Although the dynamics of
flapping-winged flight can be averaged with respect to the wingbeat period as
presented in~\cite{Schenato:flightcontrol}\cite{cheng:tandrdamping}, we develop a
linear, one-dimensional model to predict the behavior of the system, similar
to~\cite{Shigeoka:ornithopter}. In addition to the nonlinearities, the system
is underactuated due to size constraints. Many of the methods for controlling
these platforms make extensive use of GPS in outdoor
environments~\cite{kingston:timeattitude, kanade:3dvision} and expansive
motion capture systems in indoor environments. While operating indoors, using
GPS to estimate the pose of the MAV is not possible, and motion capture
requires expensive and stationary modification of the environment. The power
and payload limits on MAVs make on-board sensing difficult, so multi-agent
cooperative control schemes are apt solutions.

@@

%----------------------------------------------------------------------------%
\subsection{H$^2$Bird Ornithoper}
% humphrey
In order to conduct these experiments on cooperation, we designed a new
flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit i-Bird 
RC flyer power train and clap-fling 
wings\footnote{Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
\href{http://www.silverlit-flyingclub.com/wingsmaster/}
     {http://www.silverlit-flyingclub.com/wingsmaster/}}, 
the H$^2$Bird's wingspan is 26.5 cm and it has a flight weight of 13 grams. A 
tail propeller and servo-controlled tail elevator provide control in the yaw 
and pitch axes. The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
\href{https://github.com/biomimetics/imageproc\_pcb}
     {https://github.com/biomimetics/imageproc\_pcb}} 
electronics package consists of a 40 MIPS microprocessor, 6 DOF IMU, 
IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
lithium polymer battery. In routine flight, the H$^2$Bird averages 1.2 m/s 
ground speed, and can operate for approximately 10 minutes.
%TODO footnote for ImageProc

%----------------------------------------------------------------------------%
\subsection{Ground Station Computer Vision Platform}
% ryan
Cooperative behavior is a promising strategy for autonomous navigation of 
unexplored and unstructured environments. We intend that this system will be 
fully mobile in these environments, not tied to a static ground station or 
motion capture laboratory. To demonstrate the feasibility of this approach, 
we designed our ground station to conform to the power and weight 
constraints of highly mobile millirobotic ground vehicles, such as 
OctoRoACH, which can traverse rough terrains with speed and agility~\cite{Pullin2012Dynamic}.

@@

%----------------------------------------------------------------------------%
\subsection{Ground Station Pose Estimation}
The ground station performs simple two-dimensional pose estimation on 
H$^2$Bird using a particle filter-based motion tracking algorithm. This pose 
estimate is the feedback input to the cooperative visual servoing feedback 
loop (Sec. \ref{sec:visual_servoing_concept}). We limit our pose estimation 
approach to monocular two-dimensional object tracking in pixel space. This 
allows us to perform real-time video tracking using the modest computational 
resources available on the ground station, and similarly millirobotic ground 
vehicles.

Our pose estimation algorithm tracks H$^2$Bird motion in the frame using a 
particle filter~\cite{thrun2005probabilistic}. We initialze particles 
uniformly across the frame. To improve numerical stability, we  normalize the 
weights of all particles on each iteration, and only resample the particle 
population when the mean particle weight falls below a pre-determined 
threshold.

%............................................................................%
\subsubsection{Motion Model}
Our motion model (Eq.~\ref{eq:motion_model}) is a simple Gaussian transition 
centered around each particle's current position, augmented with an 
$\epsilon$-random uniform sampling strategy. In Eq.~\ref{eq:motion_model}, 
$x^{[i]}_t$ is the state of the $i^\text{th}$ particle at time $t$, $\Sigma$ 
is a chosen covariance matrix, and $b$ is vector of the height and width of 
the camera frame.
\begin{equation}
\label{eq:motion_model}
x^{[i]}_t \sim \begin{cases}
\mathcal{N}(x^{[i]}_{t-1}, \Sigma), &\text{with probability } 1-\epsilon \\
\mathcal{U}(0,b),& \text{with probability } \epsilon \\
\end{cases}\\
\end{equation}
The experiments presented in this paper used a diagonal covariance matrix 
with variance 256, and an $\epsilon$ of 0.3. We find that the Gaussian 
transition model is sufficient for tracking motion in which frame-to-frame 
position changes are relatively small. The $\epsilon$-random uniform 
transitions increase tracking robustness, by forcing the filter to always 
sample all of the pixel space. This prevents overconfidence and collapse, 
and decreases reacquisition delay.

%............................................................................%
\subsubsection{Emission Model}
Our emission model (Eq.~\ref{eq:emission_model}) effectively performs \naive

%----------------------------------------------------------------------------%
\subsection{Cooperative Visual Servoing}
\label{sec:visual_servoing_concept}

@@

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/pf_screencap.pdf}
\caption{Frame sequence from video feed used for tracking. Particles 
are shown in \textcolor{red}{red}, the current state estimate as a 
\textcolor{yellow}{yellow} circle, the window bounding box in
\textcolor{blue}{blue}, and the target location as a \textcolor{green}{green} 
circle. 1.~Particles initialized uniformly across the frame; 2.~Human operator 
dragging bounding box across the window; 3.~H$^2$Bird enters the frame. 
Particles converge on its location; 4.~Release H$^2$Bird, which 
begins cooperative autonomous flight toward the target; 5.~Visual servoing 
controller overshoots the window on the left, and commands a right-turn 
maneuver to recover; 6.~H$^2$Bird successfully navigates through the window.}
\label{fig:pf_screencap}
\end{figure}

@@

To model the yaw motion of the ornithopter, we developed a one-dimensional model using the physical heading angle as the output and the heading angle set by the ground station as the input. This model accounts for both the yaw dynamics of the ornithopter and the internal PID controller on the actuator inputs for the tail rotor. We decided upon a one-dimensional model to reduce the complexity of the overall model and enable simple angle inputs and outputs, which is similar to our methods during experimentation. The format of the model we decided up on is shown in Equation~\ref{eq:plain_model}.

@@

We determined the turning speed and turning radius of the H$^2$Bird
experimentally using the response of the system to the step input of a
clockwise 90 degree turn. To measure the response of the system, we recorded
the angular position data of the ornithopter calculated from the on-board
gyroscope data. We then used
MATLAB\footnote{The MathWorks, Inc. Matlab:
\href{http://www.mathworks.com/products/matlab/}
     {http://www.mathworks.com/products/matlab/}} to fit a
simple low-order model to the step response as discussed in Sec.~\ref{sec:model_attitude}.

@@

The results of the step response experiment are described in
Fig.~\ref{fig:step_response}. The first-order model that we fit to the
response is in Equation~\ref{eq:transfer_func}.

\begin{equation}
\label{eq:transfer_func}
H(s) = \frac{0.95}{1+0.62s}
\end{equation}

@@

@@

%............................................................................%
\subsubsection{Cooperative Visual Servoing}
\label{sec:visual_servoing}

@@

The results of our experimental trials are shown in
Figs.~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}. For
initial conditions in the middle of the testing space and in front of the
camera, we achieved a success rate of 80\%. This case is depicted in
Fig.~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
feasible region in Fig.~\ref{fig:flight_paths}, the success rate diminishes
to 50\% because of several factors. On the edges of the camera frame, the yaw
control inputs are large in magnitude, although this is not necessary in all
cases. Since the control algorithm we are using has no notion of depth, it
applies the same control input whether the ornithopter is close to the camera,
where small changes in heading are adequate, or far away from the camera,
where large changes in heading are necessary. Due to this phenomenon, the
controller often over- or under-compensates in certain regions, depending upon
the PID tuning and the distance from the camera. Tracking noise can also cause
passage negotiation to fail. Variations in the H$^2$Bird hand launch can cause
the system to respond differently for similar initial conditions. In addition,
the H$^2$Bird battery was changed approximately every 5 tests, adding more
variation to the H$^2$Bird performance by affecting the motor power output.

@@

We simulated the motion of the ornithopter beginning from certain positional
and angular initial conditions using several components representing the
experimental layout of our system described in Fig.~\ref{fig:block_diagram}.
For the "ornithopter dynamics" block, we used the translational model
described by~\ref{eq:total_model_plain} with the angular position model
described by~\ref{eq:transfer_func}. We tuned the PID controller for the
simulation so that paths from known positions looked similar to the behavior
we observed in the experimental data. We also inlcuded the camera capture and
pose estimation delays, and the PID controller sampling rate in the
simulation.

We determine the backwards reachable set of initial states for successful
narrow passage traversal both geometrically and in simulation
(Fig.~\ref{fig:feasible_set}). For the initial conditions, we use a uniform
grid of 0.05 meter increments within the camera viewing region, and an initial
angular position perpendicular to the narrow passage plane. The geometrically
infeasible region is computed using the minimum turning radius and the
geometry of the camera's field-of-view and is indicated in blue. Both the
regions in red and blue indicate the infeasible region determined in
simulation. The feasible region is highlighted in green. A success consisted
of a path that intersects the window at some point and does not leave the
camera viewing triangle. Fig.~\ref{fig:feasible_set} is validated by our
experimental data, as there are no successes within the geometrically
infeasible region. There are a few successes near the boundary of the
infeasible region determined in simulation, but this phenomenon can be
explained by the results in Fig.~\ref{fig:heat_map}.

Fig.~\ref{fig:heat_map} depicts the results of a Monte Carlo method of
simulating possible paths of the H$^2$Bird during our experiments. For each
point within a 0.2 meter grid within the camera viewing triangle, we simulated
10 trials using an initial heading randomly sampled from a uniform
distribution between -90$^{\circ}$ and 90$^{\circ}$. As shown in the figure,
there is a region in the middle with a 100\% probability of success, which is
a subset of the feasible region presented in Fig.~\ref{fig:feasible_set}.
Along the edges of the camera viewing triangle, however, the probability of
success decreases.

The results of our experimental trials verify the results of the Monte Carlo
simulation of the model that we used. The trials that we conducted in
Fig.~\ref{fig:flight_paths_feasible} correspond to the circled region in
Fig.~\ref{fig:heat_map}. The 80\% success rate that we experienced in the
experiments corroborates our calculated probability of success of
approximately 80\%. Additionally the diminished success rate of 50\% near the edges of the viewing triangle that we determined through experimentation supports the reduced probability of success calculated by the simulation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future Work}
%TODO 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
The authors would like to thank Fernando Garcia Bermudez for his 
assistance with the Vicon motion capture system, Andrew Pullin for his 
help with robot photography, and the members of the Biomimetic 
Millisystems Laboratory and the EECS community at the University of 
California, Berkeley for their advice and support.

>>>>

%----------------------------------------------------------------------------%
% PACKAGES
\usepackage{color}
\usepackage[pdftex]{hyperref}

@@

We also introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
features clap and fling wings, improves upon a previous design by 
utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a 
payload of 2.8 grams. We augment the ornithopter's built-in 
gyroscope-based control with a ground station, which has power and weight 
requirements appropriate for deployment on 30 gram a ground vehicle. The ground 
station provides heading estimates to the ornithoper by running a 
real-time motion tracking algorithm on a live video stream. This allows 
the ornithopter to reliably negotiate flight through narrow openings.

@@

Autonomous control of flapping winged MAVs is a complex undertaking because of
the non-linear flight
dynamics~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. One approach to
simplify these dynamic models is to use averaging over the wingbeat
period~\cite{Schenato:flightcontrol}\cite{cheng:tandrdamping}. We desire a
simpler approach, so we develop a linear, one-dimensional model to predict the
yaw behavior of the system, similar to the model that Shigeoka used
in~\cite{Shigeoka:ornithopter} for simulating the velocity of a MAV in
one-dimensionally. In addition to the nonlinearities, the system is
underactuated (there is no roll control) due to size constraints.

Many of the methods for
controlling these platforms make extensive use of GPS in outdoor
environments and expansive
motion capture systems in indoor environments~\cite{kingston:timeattitude}\cite{kanade:3dvision}. While operating indoors, using
GPS to estimate the pose of the MAV is not possible, and motion capture
requires expensive and stationary modification of the environment. The power
and payload limits on MAVs make on-board sensing difficult, so multi-agent
cooperative control schemes are apt solutions.

@@

%----------------------------------------------------------------------------%
\subsection{H$^2$Bird Ornithoper}
% humphrey
In order to conduct these experiments on cooperation, we designed a new
flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit I-Bird 
RC flyer power train and clap-fling 
wings\footnote{\url{http://www.silverlit-flyingclub.com/wingsmaster/}}, 
the H$^2$Bird's wingspan is 26.5 cm and it has a flight weight of 13 grams. A 
tail propeller and servo-controlled tail elevator provide control in the yaw 
and pitch axes. The on-board ImageProc 2.4\footnote{Expand! Accredit it to Stan Baek!} 
electronics package consists of a 40 MIPS microprocessor, 6 DOF IMU, 
IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
lithium polymer battery. In routine flight, the H$^2$Bird averages 1.2 m/s 
ground speed, and can operate for approximately 10 minutes.
%TODO footnote for ImageProc

%----------------------------------------------------------------------------%
\subsection{Ground Station Computer Vision Platform}
% ryan
Cooperative behavior is a promising strategy for autonomous navigation of 
unexplored and unstructured environments. We intend that this system will be 
fully mobile in these environments, not tied to a static ground station or 
motion capture laboratory. To demonstrate the feasibility of this approach, 
we designed our ground station to conform to the power and weight 
constraints of highly mobile millirobotic ground vehicles, such as 
OctoRoACH, which can traverse rough terrains with speed and agility [CITATION]. 
%TODO OctoRoACH citation

@@

%----------------------------------------------------------------------------%
\subsection{Ground Station Pose Estimation}
%TODO ryan
The ground station performs simple two-dimensional pose estimation on the H$^2$Bird using 
a particle filter-based motion tracking algorithm. This pose estimate is 
the feedback input to the cooperative visual servoing feedback loop (Sec. \ref{sec:visual_servoing_concept}).

We limit our pose estimation approach to monocular two-dimensional object 
tracking in pixel-space. This allows us to perform high-rate video tracking 
using the modest computational resources available on the ground station, 
and similarly millirobotic ground vehicles.

~\cite{thrun2005probabilistic}

%----------------------------------------------------------------------------%
\subsection{Cooperative Visual Servoing}
\label{sec:visual_servoing_concept}
%TODO ryan

@@

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/pf_screencap.pdf}
\caption{Frame sequence from video feed used for tracking. Particles 
are shown in \textcolor{red}{red}, the current state estimate as a 
\textcolor{blue}{blue} circle, the window bounding box in
\textcolor{red}{red}, and the target location as a \textcolor{green}{green} 
circle. 1.~Particles initialized uniformly across the frame; 2.~Human operator 
dragging bounding box across the window; 3.~H$^2$Bird enters the frame. 
Particles converge on its location; 4.~Release H$^2$Bird, which 
begins cooperative autonomous flight toward the target; 5.~Visual servoing 
controller overshoots the window on the left, and commands a right-turn 
maneuver to recover; 6.~H$^2$Bird successfully navigates through the window.}
\label{fig:pf_screencap}
\end{figure}

@@

To model the yaw motion of the ornithopter, we developed a one-dimensional model using the physical heading angle as the output and the heading angle set by the ground station as the input. This model accounts for both the yaw dynamics of the ornithopter and the internal PID controller on the actuator inputs for the tail rotor. We decided upon a one-dimensional model to reduce the complexity of the overall model and enable simple angle inputs and outputs, which is similar to our methods during experiementation. The format of the model we decided up on is shown in Equation~\ref{eq:plain_model}.

@@

We determined the turning speed and turning radius of the H$^2$Bird
experimentally using the response of the system to the step input of a
clockwise 90 degree turn. To measure the response of the system, we recorded
the angular position data of the ornithopter calculated from the on-board
gyroscope data. We then used
MATLAB\footnote{\url{http://www.mathworks.com/products/matlab/}} to fit a
simple low-order model to the step response as discussed in Section~\ref{sec:model_attitude}.

@@

The results of the step response experiment are described in
Figure~\ref{fig:step_response}. The first-order model that we fit to the
response is in Equation~\ref{eq:transfer_func}.

\begin{equation}
\label{eq:transfer_func}
H(s) = \frac{0.95}{1+0.62s}
\end{equation}
%TODO H(s) in a labeled equation
%TODO footnote for MATLAB

@@

%TODO equation(s) for how you arrive at the minimum turning radius
%TODO footnote for MATLAB

@@

%............................................................................%
\subsubsection{Cooperative Visual Servoing}
\label{sec:visual_servoing}
%TODO cameron

%TODO make these span the page, with identical scales/style

@@

The results of our experimental trials are shown in
Figures~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}. For
initial conditions in the middle of the testing space and in front of the
camera, we achieved a success rate of 80\%. This case is depicted in
Figure~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
feasible region in Figure~\ref{fig:flight_paths}, the success rate diminishes
to 50\% because of several factors. On the edges of the camera frame, the yaw
control inputs are large in magnitude, although this is not necessary in all
cases. Since the control algorithm we are using has no notion of depth, it
applies the same control input whether the ornithopter is close to the camera,
where small changes in heading are adequate, or far away from the camera,
where large changes in heading are necessary. Due to this phenomenon, the
controller often over- or under-compensates in certain regions, depending upon
the PID tuning and the distance from the camera. Tracking noise can also cause
passage negotiation to fail. Variations in the H$^2$Bird hand launch can cause
the system to respond differently for similar initial conditions. In addition,
the H$^2$Bird battery was changed approximately every 5 tests, adding more
variation to the H$^2$Bird performance by affecting the motor power output.

@@

We simulated the motion of the ornithopter beginning from certain positional
and angular initial conditions using several components representing the
experimental layout of our system described in Figure~\ref{fig:block_diagram}.
For the "ornithopter dynamics" block, we used the translational model
described by~\ref{eq:total_model_plain} with the angular position model
described by~\ref{eq:transfer_func}. We tuned the PID controller for the
simulation so that paths from known positions looked similar to the behavior
we observed in the experimental data. We also inlcuded the camera capture and
pose estimation delays, and the PID controller sampling rate in the
simulation.

We determine the backwards reachable set of initial states for successful
narrow passage traversal both geometrically and in simulation
(Fig.~\ref{fig:feasible_set}). For the initial conditions, we use a uniform
grid of 0.05 meter increments within the camera viewing region, and an initial
angular position perpendicular to the narrow passage plane. The geometrically
infeasible region is computed using the minimum turning radius and the
geometry of the camera's field-of-view and is indicated in blue. Both the
regions in red and blue indicate the infeasible region determined in
simulation. The feasible region is highlighted in green. A success consisted
of a path that intersects the window at some point and does not leave the
camera viewing triangle. Figure~\ref{fig:feasible_set} is validated by our
experimenal data, as there are no successes within the geometrically
infeasible region. There are a few successes near the boundary of the
infeasible region determined in simulation, but this phenomenon canbe
explained by the results in Figure~\ref{fig:heat_map}.

Figure~\ref{fig:heat_map} depicts the results of a Monte Carlo method of
simulating possible paths of the H$^2$Bird during our experiments. For each
point within a 0.2 meter grid within the camera viewing triangle, we simulated
10 trials using an initial heading randomly sampled from a uniform
distribution between -90$^{\circ}$ and 90$^{\circ}$. As shown in the figure,
there is a region in the middle with a 100\% probability of success, which is
a subset of the feasible region presented in Figure~\ref{fig:feasible_set}.
Along the edges of the camera viewing triangle, however, the probability of
success decreases.

The results of our experimental trials verify the results of the Monte Carlo
simulation of the model that we used. The trials that we conducted in
Figure~\ref{fig:flight_paths_feasible} correspond to the circled region in
Figure~\ref{heat_map}. The 80\% success rate that we experienced in the
experiments corroborates our calculated probability of success of
approximately 80\%. Additionally, the diminished success rate of 50\% near the
edges of the viewing triangle that we determined through experimentation
supports the reduced probability of success calculated by the simulation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future Work}
%TODO ryan
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgements}
The authors would like to thank Fernando Garcia Bermudez for his 
assistance with the Vicon motion capture system, Andrew Pullin for his 
help with robot photography, and the members of the Biomimetic 
Millisystems Laboratory and the EECS community at the University of 
California, Berkeley for their advice and support.
